---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Load necessary packages

```{r}
library(ggplot2)
library(scales)
library(Hmisc)
```

## Loading and preprocessing the data

Downlaod and unzip the data.

```{r}
FILE_URL <- 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'
FILE_NAME_ZIP <- 'activity.zip'
FILE_NAME_CSV <- 'activity.csv'

if (!file.exists(FILE_NAME_CSV)) {
  download.file(FILE_URL, destfile=FILE_NAME_ZIP)
  unzip(FILE_NAME_ZIP)
}

activities = read.csv(FILE_NAME_CSV)
```

Get summary of the data.

```{r}
summary(activities)
```

Notice we have some NAs in `steps`.

## What is mean total number of steps taken per day?

Calculate sum of steps group by dates.

```{r}
activitiesStepsByDate <- aggregate(
  activities[, c('steps')],
  by=list(activities$date),
  sum,
  na.rm=TRUE)
names(activitiesStepsByDate) <- c('date', 'stepsSum')
head(activitiesStepsByDate)
```

Histogram of the total number of steps taken by date. (I am not sure if histogram is a right plot here because histogram is meant for data has continuous x values, but dates are actually not.)

```{r}
ggplot(data=activitiesStepsByDate,
       aes(activitiesStepsByDate$date, activitiesStepsByDate$stepsSum)) +
  stat_summary(fun.y=sum, geom='bar') +
  theme(axis.text.x = element_text(angle=90, hjust=1))
```

Calculate and report the mean and median total number of steps taken per day.

```{r}
activitiesStepsByDateMean = mean(activitiesStepsByDate$stepsSum)
activitiesStepsByDateMid = median(activitiesStepsByDate$stepsSum)
```

* Mean number of steps taken per day: `r activitiesStepsByDateMean`
* Median number of steps taken per day: `r activitiesStepsByDateMid`

## What is the average daily activity pattern?

```{r}
activitiesStepsByInterval <- aggregate(
  x=list(meanSteps=activities$steps),
  by=list(interval=activities$interval),
  FUN=mean,
  na.rm=TRUE
) 
```

Make a time series plot.

```{r}
ggplot(data=activitiesStepsByInterval, aes(x=interval, y=meanSteps)) +
  geom_line() +
  xlab('5-minute interval in a day') +
  ylab('Average steps')
```

Find the interval with maximum average steps.

```{r}
maxStepsInterval <- which.max(activitiesStepsByInterval$meanSteps)
```

It's the `r maxStepsInterval`th interval.

## Imputing missing values

```{r}
countNaSteps = length(which(is.na(activities$steps)))
```

There are `r countNaSteps` NA steps values.

Fill NA values with mean number.

```{r}
activitiesImputed <- activities
activitiesImputed$steps <- impute(activitiesImputed$steps, fun=mean)
```

Plot the histogram.

```{r}
activitiesStepsByDateImputed <- aggregate(
  activitiesImputed[, c('steps')],
  by=list(activitiesImputed$date),
  sum,
  na.rm=TRUE)
names(activitiesStepsByDateImputed) <- c('date', 'stepsSum')

ggplot(data=activitiesStepsByDateImputed,
       aes(date, stepsSum)) +
  stat_summary(fun.y=sum, geom='bar') +
  theme(axis.text.x = element_text(angle=90, hjust=1))
```

Calculate mean and median.

```{r}
activitiesStepsByDateMeanImputed = mean(activitiesStepsByDateImputed$stepsSum)
activitiesStepsByDateMidImputed = median(activitiesStepsByDateImputed$stepsSum)
```

* Mean number of steps taken per day (after imputing): `r activitiesStepsByDateMeanImputed`
* Median number of steps taken per day (after imputing): `r activitiesStepsByDateMidImputed`

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
activitiesImputed$dateType <- ifelse(as.POSIXlt(activitiesImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

Make a panel plot containing a time series plot

```{r}
averagedActivityDataImputed <- aggregate(steps ~ interval + dateType, data=activitiesImputed, mean)
ggplot(averagedActivityDataImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```