---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Load necessary packages

```{r}
library(ggplot2)
library(scales)
library(Hmisc)
```

## Loading and preprocessing the data

Downlaod and unzip the data.

```{r}
FILE_URL <- 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'
FILE_NAME_ZIP <- 'activity.zip'
FILE_NAME_CSV <- 'activity.csv'

if (!file.exists(FILE_NAME_CSV)) {
  download.file(FILE_URL, destfile=FILE_NAME_ZIP)
  unzip(FILE_NAME_ZIP)
}

activities = read.csv(FILE_NAME_CSV)
```

Get summary of the data.

```{r}
summary(activities)
```

Notice we have some NAs in `steps`.

## What is mean total number of steps taken per day?

Calculate sum of steps group by dates.

```{r}
activitiesStepsByDate <- tapply(activities$steps, activities$date, sum, na.rm=TRUE)
head(activitiesStepsByDate)
```

Histogram of the total number of steps taken by date.

```{r}

```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
